Morse Theory

### 2020-11-29 paper.tex
dassd

\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\X}{\mathbb{X}}
\newcommand{\imply}{\Rightarrow}

@@

\tableofcontents
\newpage
\section{Morse Theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % begin N1
\subsection{Introduction}
    \paragraph{\textbf{Morse Lemma.}}
        Suppose that the point $a \in \mathbb{R}^k$ is a nondegenerate
        critical point of the function $f$, and
        $$
            (h_{ij})= \left(  \frac{\partial^2f}{\partial x_i \partial x_j} (a) \right)
        $$
        is the Hessian of $f$ at $a$. Then there exists a local coordinate system
        $(x_1,\ldots,x_k)$ around $a$ such that
        $$
            f=f(a)+\sum{h_{ij}x_ix_j}
        $$
        near $a$.
        \\
        or
        $$
            f=-x_1^2-x_2^2-\ldots-x_\lambda^2+x_\lambda^2+\ldots+x_n^2+f(a)
        $$
        where $\lambda$ is the index of $f$ at $a$.

@@

     When we square $y_1$, we get
     \begin{equation}
        y_1^2=\pm h_{11}x_1^2\pm2\sum_{i=2}^{n}{x_1x_ih_{1i}}\pm\frac{\left(\sum_{i=2}^{n}{x_ih_{1i}}\right)^2}{h_{11}}\tag{1.9}
     \end{equation}
     where the signs are either positive or negative, depending on the sign of $h_{11}$. Using equation
     $1.5$, we can verify that $f$ can be expressed in the following way with respect to this
     coordinate system on the restricted domain $\tilde{N}_0$.
     \begin{equation}
        f=\pm y_1^2 + \sum_{i=2}^{n}\sum_{j=2}^{n}{x_ix_jh_{ij}}-\frac{\left(\sum_{i=2}^{n}{x_ih_{1i}}\right)^2}{h_{11}}\tag{1.10}
     \end{equation}
     where the sign of the $y_1^2$ term is positive or negative, depending on the sign of $h_{11}$. Staying
     consistent with our goals, we notice that the first term is in the standard quadratic form seen
     in the Morse Lemma formulation, whereas the rest of the terms depend on local coordinates
     $x_i$ whereby $i\not=k \ (k=1)$. By induction from $k=1$ to $k=n$, we prove the Morse Lemma. $\blacksquare$
% % end N1
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % begin M1

\subsection{Proof that$f: S^2 \to \R, f(x,y,z) = z$ is a Morse function}
\textbf{Ex.} Show that the function $f: S^2 \to \R, f(x,y,z) = z$ is a Morse function. \\

@@

\subsection{Examples of Morse functions} 

@@

\subsection{Examples of non-Morse functions} 

@@

\newpage
\section{Reeb graphs}
     \subsection{Introduction to Reeb graphs}
     \paragraph{}
      Suppose we have an equivalence relation, $\sim$, defined on a topological space $\X$. Let $\X_\sim$
      be the set of equivalence classes and let $\phi : \X \to \X_\sim$ map each point $x$ to its equivalence class.

     \paragraph{\textbf{Definition 1.2.1}}
      Let $f: \X \to \R$ be continuous and call a component of a level set a contour. Two
      points $x,y \in \X$ are equivalent if they belong to the same component of $f^{-1}(t)$
      with $t=f(x)=f(y)$. The Reeb graph of $f$, denoted as $R(f)=\X_\sim$, is the quotient space 
      defined by this equivalence relation.

      By construction, the Reeb graph has a point for each contour and the connection is 
      provided by the map $\phi:\X\to R(f)$. Letting $\pi:R(f)\to\R$ be defined such that $f(x)=\pi(\phi(x))$
      we can construct the level set by going backward, from the real line to the Reeb graph to the 
      topological space. Given $t\in\R$ we get $\pi^{-1}(t)$, a collection of points in $R(f)$,
      and $\phi^{-1}(\pi^{-1}(t))$, the corresponding collection of contours that make up the level set defined by $t$.

      \paragraph{Notes:}We have a continuous surjection $\phi:\X\to R(f)$ which maps components 
      to components. Furthermore, a loop in $\X$ that maps to a loop in $R(f)$ is not contractible
      and two loops in $\X$ that map to different loops in $R(f)$ are not homologous. It follows
      that the number of components is preserved and the number of loops cannot increase:
      $$
      \beta_0(R(f))=\beta_0(\X)
      $$
      $$
      \beta_1(R(f))\le\beta_1(\X)
      $$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % begin N2 
% % end N2
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\textbf{Lamma ?}\\

@@

>>>>

\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\imply}{\Rightarrow}

@@

\tableofcontents
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % begin N1
    \paragraph{\textbf{Morse Lemma.}}
        Suppose that the point $a \in \mathbb{R}^k$ is a nondegenerate
        critical point of the function $f$, and
        $$
            (h_{ij})= \left(  \frac{\partial^2f}{\partial x_i \partial x_j} (a) \right)
        $$
        is the Hessian of $f$ at $a$. Then there exists a local coordinate system
        $(x_1,\ldots,x_k)$ around $a$ such that
        $$
            f=f(a)+\sum{h_{ij}x_ix_j}
        $$
        near $a$.
        \\
        or
        $$
            f=-x_1^2-x_2^2-\ldots-x_\lambda^2+x_\lambda^2+\ldots+x_n^2+f(a)
        $$
        where $\lambda$ is the index of $f$ at $a$.

@@

     When we square $y_1$, we get
     \begin{equation}
        y_1^2=\pm h_{11}x_1^2\pm2\sum_{i=2}^{n}{x_1x_ih_{1i}}\pm\frac{\left(\sum_{i=2}^{n}{x_ih_{1i}}\right)^2}{h_{11}}\tag{1.9}
     \end{equation}
     where the signs are either positive or negative, depending on the sign of $h_{11}$. Using equation
     $1.5$, we can verify that $f$ can be expressed in the following way with respect to this
     coordinate system on the restricted domain $\tilde{N}_0$.
     \begin{equation}
        f=\pm y_1^2 + \sum_{i=2}^{n}\sum_{j=2}^{n}{x_ix_jh_{ij}}-\frac{\left(\sum_{i=2}^{n}{x_ih_{1i}}\right)^2}{h_{11}}\tag{1.10}
     \end{equation}
     where the sign of the $y_1^2$ term is positive or negative, depending on the sign of $h_{11}$. Staying
     consistent with our goals, we notice that the first term is in the standard quadratic form seen
     in the Morse Lemma formulation, whereas the rest of the terms depend on local coordinates
     $x_i$ whereby $i\not=k \ (k=1)$. By induction from $k=1$ to $k=n$, we prove the Morse Lemma. $\blacksquare$
% % end N1
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % begin M1
\textbf{Ex.} Show that the function $f: S^2 \to \R, f(x,y,z) = z$ is a Morse function. \\

@@

Examples of Morse functions: \\

@@

Examples of functions that are not Morse: \\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % begin N2 
% % end N2
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{lemma}

@@

\end{lemma}
